---
title: "Elution Time of FPLC"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = F)
pkgs <- c("dplyr", "ggplot2", "tidyr", "DT")
for (pkg in pkgs) {
        suppressPackageStartupMessages(library(pkg, character.only = T))
}

# data <- readRDS("../data/chromatogram1.rds")
data2 <- readRDS("../data/chromatogram2.rds")
```

## The chromatogram

Figure 1 is the chromatogram of six ERCC samples.

```{r}
ggplot(data2, aes(x = ml, y = mAU)) +
        geom_line(aes(group = sample, color = sample)) +
        labs(title = "Figure 1. The chromatogram of six ERCC samples",
             x = "Volume eluted (ml)", y = "UV readings (mAu)") +
        theme_bw()
```

## The variation of the site of compound peaks

Table 1 showed the mean and the standard deviation of the site of compound peaks
based on the volume eluted. The peaks from left to right corresponds to 
LDL, HDL and albumin. The units of mean and sd are **ml**.

```{r, message=FALSE}

# Identidy the fraction regions
peak_ldl <- vector("numeric", length = length(levels(data2$sample)))
peak_hdl <- vector("double", length = length(levels(data2$sample)))
peak_alb <- vector("double", length = length(levels(data2$sample)))
for (i in seq_along(levels(data2$sample))){
        ldl <- filter(data2, sample == levels(data2$sample)[i] & ml<10) %>%
                filter(mAU == max(mAU))
        peak_ldl[i] <- as.numeric(ldl[1,1])
        hdl <- filter(data2, sample == levels(data2$sample)[i], ml<13 & ml>10) %>%
                filter(mAU == max(mAU))
        peak_hdl[i] <- as.numeric(hdl[1,1])
        alb <- filter(data2, sample == levels(data2$sample)[i], ml>13) %>%
                filter(mAU == max(mAU))
        peak_alb[i] <- as.numeric(alb[1,1])
}


peaks <- data.frame(
        sample = levels(data2$sample),
        ldl = peak_ldl,
        hdl = peak_hdl,
        alb = peak_alb
) %>%
        pivot_longer(cols = 2:4, names_to = "comp", "values_to" = "ml") %>%
        mutate(comp = factor(comp, 
                               levels = c("ldl", "hdl", "alb"), 
                               labels = c("LDL", "HDL", "Albumin")))
summary <- peaks %>%
        group_by(comp) %>%
        summarise(
                mean = mean(ml),
                sd = sd(ml)
        ) %>%
        arrange(comp)
datatable(summary, rownames = F) %>%
        formatSignif(columns = 2:3, digits = 3)
```

Figure 2 showed the mean and standard deviation of these three compoments.
```{r}
ggplot(peaks, aes(x = comp, y = ml)) +
        geom_jitter(aes(color = sample), width = 0.1) +
        stat_summary(fun.data=mean_sdl, fun.args = list(mult=1), 
                     geom="errorbar", color="red", width=0.1) +
        labs(title = "Figure 2. The mean and standard deviation of elution time",
             x = "compoments", y = "Volume eluted (ml)") +
        theme_bw()
```

Figure 3 added the lines of the average volume eluted of three compoments to the chromatogram.
```{r}
ggplot(data2, aes(x = ml, y = mAU)) +
        geom_line(aes(group = sample, color = sample)) +
        geom_vline(xintercept = summary$mean, linetype = 2) +
        labs(title = "Figure 3. The chromatogram of six ERCC samples \n with the average volume eluted of three compoments",
             x = "Volume eluted (ml)", y = "UV readings (mAu)") +
        theme_bw()
```

