---
title: "Untargeted HDL lipidomics peak height"
author: "Cynthia"
format:
  html:
    self-contained-math: true
    toc: true
    toc-location: left
server: shiny
---

## *CVs of pooled samples*

*The CVs of pooled samples for each lipid species were calculated using the following formula:*

$$
CV_{lipid\ species} = SD_{lipid\ species}/Mean_{lipid\ species} * 100\%
$$

*We are less confident with the lipid species with a high CV. These lipid species will be excluded for further analysis.*

```{r, cv}
plotOutput("hist")
verbatimTextOutput("cv30")
```


Select lipid species without iSTD and normalize their peak heights to total peak heights within each sample.

Total peak heights consists of major lipid species which can be quantified with iSTDs and minor species which don't have corresponding iSTDs. The percentage of both types of species are shown below.

```{r, outputPie, fig.height=6, fig.width=10}
plotOutput("pie")
```


## Barplot


```{r, outputUI}
lipidNames <- readRDS("lipidNames.rds")
selectInput("feature", "Lipid species", choices = lipidNames)
```



```{r, outputPlot}
plotOutput("bars")
```

